% \section{To Do}
% \label{s:todo}

% \finish{discuss multiple equations/contractions} 

% The formalisation of multi-variable context (or expressions) and the version of
% ``unique solutions'' theorems with multiple equations/contractions, is
% a relative hard problem. To the best of our knowledge, so far there's no formalization (in any kind of
% process calculi) touching these areas yet. However, we've made a few initial steps
% of this interesting problem.

\section{Conclusions and future work}
\label{s:concl}

We have highlighted a formalisation of the theory of CCS in the 
HOL4 theorem prover (for lack of space we have not discussed 
the formalisation of some basic algebraic theory, of the basic
properties of the expansion preorder,   and of a few
 versions of `bisimulation up to'
techniques). % such as  bisimulation up-to bisimilarity). 
The formalisation has focused on the theory of
unique solution of equations and contractions. 
It has also allowed us to further develop the theory,
notably the basic properties of rooted contraction, and the unique
solution theorem for it with respect to rooted bisimilarity. 
The formalisation brings up and exploits similarities between results
and proofs for different equivalences and preorders. 
We think that the statements in the formalisation are easy to read and
understand, as they are very close to the orginal statements found in
standard CCS textbooks \cite{Gorrieri:2015jt,Mil89}.

For the future work, it would be worth extending the current work
to multi-variable equations/contractions. \hl{Preliminary work shows that,
the key idea could be using unguarded constants as free variables
(\texttt{FV}) and define guardness directly on type CCS (instead of CCS
$\rightarrow$ CCS) then connect to previous context:}
\begin{alltt}
\HOLTokenTurnstile{} \HOLConst{weakly_guarded1} \HOLFreeVar{E} \HOLSymConst{\HOLTokenEquiv{}}
   \HOLSymConst{\HOLTokenForall{}}\HOLBoundVar{X}. \HOLBoundVar{X} \HOLSymConst{\HOLTokenIn{}} \HOLConst{FV} \HOLFreeVar{E} \HOLSymConst{\HOLTokenImp{}} \HOLSymConst{\HOLTokenForall{}}\HOLBoundVar{e}. \HOLConst{CONTEXT} \HOLBoundVar{e} \HOLSymConst{\HOLTokenConj{}} (\HOLBoundVar{e} (\HOLConst{var} \HOLBoundVar{X}) \HOLSymConst{=} \HOLFreeVar{E}) \HOLSymConst{\HOLTokenImp{}} \HOLConst{WG} \HOLBoundVar{e}
\end{alltt}

Formalising other equivalences and preorders could also be considered,
notably the trace equivalences.
On the other side, one could examine the formalisation of a different
approach \cite{DurierHS17} to unique
solutions, in which the use of contraction is
replaced by semantic conditions on process transitions such as
divergence. \hl{The problem of formalising value-passing CCS in a
more elegant way with minimal efforts (by exploiting the type variable
of actions, for example) also deserves some thinking.}

% Further plan on the formalisation mainly includes: 1) the extension to
% multi-variable equation/contractions. 2) the support of recursion
% operators in 
% CCS context and expressions.

% We believe that, beside the discovery of the rooted contraction $\rcontr$
% with a more elegant unique solution theorem,
% our CCS formalisation in HOL4 has also
% provided a solid formal framework for future theoretical developments in
% Concurrency Theory, particularily for process algebras like CCS. It's
% easily understandable, with statements extremely close to the original
% textbook. The logic foundations of HOL makes the whole work (or individual
% parts) easily portable to other theorem provers.

\paragraph{Acknowledgements}

We have benefitted from suggestions and comments 
from several people from the HOL
community, including Robert Beers, Jeremy Dawson,
Ramana Kumar,
Michael Norrish, 
Konrad Slind, and
Thomas Tuerk.
%
The second half of this
paper was written in memory of Mike J. Gordon, the creator of HOL theorem prover.
