%%%% -*- Mode: LaTeX -*-
%%
%% This is the draft of the 2nd part of EXPRESS/SOS 2018 paper, co-authored by
%% Prof. Davide Sangiorgi and Chun Tian.

\documentclass[submission]{eptcs} % required template (EPTCS)
\providecommand{\event}{EXPRESS/SOS 2018} % Name of the event you are submitting to

\usepackage[T1]{fontenc}
\usepackage{lmodern} % Latin Modern fonts (better than CM-Super)
\usepackage[utf8]{inputenc} % no need for XeTeX, which always uses UTF-8

%% Math symbol packages
\usepackage{amsmath}
\usepackage{amssymb}
\usepackage{mathrsfs}

\usepackage{stmaryrd}
\SetSymbolFont{stmry}{bold}{U}{stmry}{m}{n}
% \cupdot (the best one)
\newcommand{\cupdot}{\mathbin{\mathaccent\cdot\cup}}

\usepackage{amsthm}

\newtheorem{definition}{Definition}[section]
\newtheorem{example}[definition]{Example}
\newtheorem{lemma}[definition]{Lemma}
\newtheorem{theorem}[definition]{Theorem}
\newtheorem{corollary}[definition]{Corollary}
\newtheorem{proposition}[definition]{Proposition} 
\newtheorem{remark}[definition]{Remark}

% HOL theorem embedding support
\usepackage{holindex}
\usepackage{alltt}
\include{tokens}

\usepackage{listings}
\renewcommand{\ttdefault}{cmtt} % {pcr}
\lstset{tabsize=8,language=ML,basicstyle=\small\ttfamily\bfseries}

%
% LaTeX math formula spacing:
%
% \enskip	leave a horizontal space of respectively half an em
% \quad		space equal to the current font size (= 18 mu), one em
% \qquad	twice of \quad (= 36 mu), i.e. two ems
% \enspace	it's inherited from Plain TeX and is almost the same
%                      as \enskip, but technically it is a kern, rather than a skip.
% \,	3/18 of \quad (= 3 mu)
% \:	4/18 of \quad (= 4 mu)
% \;	5/18 of \quad (= 5 mu)
% \!	-3/18 of \quad (= -3 mu)
% \ (space after backslash!) equivalent of space in normal text

\usepackage{graphicx}
\usepackage[all,cmtip]{xy}

\input{commonmacros.sty}
\input{davidemacro.sty}
\usepackage{DSarrow} % this is something to use extensible arrows in transitions

% \usepackage{pgf,tikz}
% \usetikzlibrary{arrows}

\title{The coarsest precongruence contained in bisimilarity 
contraction and its unique solution theorem}
\author{Davide Sangiorgi
\institute{Universit\`a di Bologna and INRIA\\Bologna, Italy}
\email{davide.sangiorgi@unibo.it}
\and Chun Tian
\institute{Fondazione Bruno Kessler\thanks{Part of this work was
    carried out when the author was studying in University of
    Bologna.}\\Trento, Italy}
\email{ctian@fbk.eu}
}
\def\titlerunning{Coarsest precongruence contained in bisimilarity contraction}
\def\authorrunning{D. Sangiorgi \& C. Tian}

\begin{document}
\maketitle

\begin{abstract}
Milner's ``unique solution of equations'' theorem for (rooted) weak
bisimilarity have severe syntactical limitations, i.e.\;the
equations must be both strongly guarded and sequential. By replacing
equations with special inequations called ``contraction'', Sangiorgi
has proven the ``unique solution of contractions'' theorem for weak
bisimialarity, which requires only weakly guarded equations. However,
contraction is not (pre)congruence under direct sums of processes.
To overcome this difficulty, we further moved to ``rooted
contraction'', which is the coarsest precongruence contained in
contraction bisimilarity. Using rooted contraction one obtains an
unique-solution theorem that is valid for 
\emph{rooted bisimilarity} (hence also for bisimilarity itself) while
requires true weak guardness (with direct sums). All results were
formalized in HOL theorem prover (HOL4), as part of a rather complete
formalization of process algebra CCS. We show some highlights of this
formalization work.
\end{abstract}

% part 1 (Sangiorgi)
\input{intro}
\input{ccs}
\input{expansion}
\input{equa}
\input{contraction}
\input{unique}

% part 2 (Tian)
\input{formal}
\input{sos}
\input{bisim}

\input{part2}

\input{related}
\input{conclusions}

\bibliographystyle{eptcs}
\bibliography{generic}
\end{document}
